---
title: "Data Integrity Check"
output: html_document
---

### Purpose

This document compares monthly counts at various stages of the model process.  We show how many food inspection records (and critical violations) there were

1. In the original download, 
2. After filtering, and
2. After taking out any NA's (which indicates that a data sources wasn't able to merged into the food inspection data).

```{r, echo=FALSE, warning=FALSE, results='hide', message=FALSE}

## Initialize and load data... (hidden in report output)

geneorama::set_project_dir("food-inspections-evaluation")
##==============================================================================
## INITIALIZE
##==============================================================================
## Remove all objects; perform garbage collection
rm(list=ls())
gc(reset=TRUE)
## Detach libraries that are not used
geneorama::detach_nonstandard_packages()
## Load libraries that are used
geneorama::loadinstall_libraries(c("data.table", "MASS"))
## Load custom functions
geneorama::sourceDir("CODE/functions/")

##==============================================================================
## LOAD CACHED RDS FILES
##==============================================================================
## Import the key data sets used for prediction
foodInspect <- readRDS("DATA/food_inspections.Rds")
foodInspect$Violations <- NULL
violation_dat <- readRDS("DATA/violation_dat.Rds")
dat_model <- readRDS("DATA/dat_model.Rds")

##==============================================================================
## ADD FIELDS AND APPLY FILTER TO foodInspect
##==============================================================================
## Remove duplicate IDs first
foodInspect <- foodInspect[!duplicated(Inspection_ID)]

## Limit dates
foodInspect <- foodInspect[Inspection_Date > as.IDate("2011-09-01")]

## Join in the violation matrix
foodInspect <- merge(x = foodInspect, 
                     y = violation_dat, 
                     by = "Inspection_ID", 
                     all.x = TRUE)

foodInspect[ , criticalFound := pmin(1, criticalCount)]

foodInspect[i = TRUE, 
            j = Facility_Type_Clean := 
                categorize(x = Facility_Type,
                           primary = list(Restaurant = "restaurant",
                                          Grocery_Store = "grocery"),
                           ignore.case = TRUE)]
foodInspect_filtered <- filter_foodInspect(foodInspect)

##==============================================================================
## ADD FIELDS AND APPLY FILTER TO dat_model
##==============================================================================
dat_model[ , criticalFound := pmin(1, criticalCount)]
## Only keep "Retail Food Establishment"
dat_model <- dat_model[LICENSE_DESCRIPTION == "Retail Food Establishment"]
## Remove License Description
dat_model[ , LICENSE_DESCRIPTION := NULL]
dat_model <- na.omit(dat_model)

```

### Construct tables of counts, and critical violations found 

```{r}
tab_food <- foodInspect[
    i = TRUE, 
    j = list(N = .N, 
             critFound = sum(criticalFound)), 
    keyby = list(month = round(Inspection_Date, "month"))]

tab_food_open_canvass <- foodInspect[
    i = Inspection_Type == "Canvass" &
        !Results %in% c('Out of Business',
                        'Business Not Located',
                        'No Entry'), 
    j = list(N = .N, 
             critFound = sum(criticalFound)), 
    keyby = list(month = round(Inspection_Date, "month"))]

tab_food_filtered <- foodInspect_filtered[
    i = TRUE, 
    j = list(N = .N, 
             critFound = sum(criticalFound)), 
    keyby = list(month = round(Inspection_Date, "month"))]

tab_modeldat <- dat_model[
    i = TRUE, 
    j = list(N = .N, 
             critFound = sum(criticalFound)), 
    keyby = list(month = round(Inspection_Date, "month"))]
```

### Monthly counts at each step in "filtering"

```{r}
tab_food
tab_food_open_canvass
tab_food_filtered
tab_modeldat

```

### Monthly counts at each stage in "filtering", merged into a single table

```{r}
tab_all <- merge(tab_food, 
                 merge(tab_food_open_canvass, 
                       merge(tab_food_filtered,
                             tab_modeldat, 
                             all = TRUE,
                             suffixes = c(".Filtered", ".Model")), 
                       all=TRUE),
                 all=TRUE,
                 suffixes = c(".ALL", ".Open"))
tab_all
```


### Violations, tabluated by "is.na"

Generally, it looks like data that was incomplete was lower in violation counts.

```{r}
geneorama::set_project_dir("food-inspections-evaluation")

## Read in the original dat model before any filtering is applied
dat_model_orig <- readRDS("DATA/dat_model.Rds")

## Luckily the restaurants with missing business data mostly appear to have
## lower counts of critical and serious violations
geneorama::NAsummary(dat_model_orig)
dat_model_orig[ , table(is.na(Business_ID))]
dat_model_orig[
    i = TRUE,
    j = list(mean_critical = mean(criticalCount), sd_critical = sd(criticalCount),
             mean_serious = mean(seriousCount), sd_serious = sd(seriousCount),
             mean_minor = mean(minorCount), sd_minor = sd(minorCount)), 
    is.na(Business_ID)]
```











